### 2013-12-01 paper/main.tex
normalization corrections for engine model
Corrections to normalize the engine model have been found
which will accept a step input and produce a response near one.

In order to build a controller the model needs to be simplified
but the presence of two inputs complicates matters.
To resolve this issue the torque can be set zero.
With this change it can be simplified by first recognizing that
it matches the form shown in Figure \ref{fig:dd} which
has the transfer function in Equation \ref{eq:dd}.

@@

\begin{samepage}
Then substituting them in to Equation \ref{eq:dd} results in
the simplified from of Equation \ref{eq:lems}.
\begin{align}
	\frac{y}{u} &= \frac{(1.699) (8.5683) (2.98) (1)}
		{(z^2 - 0.9025z)(z - 0.9354)(1) + (8.5683)(2.98)(0.00093)} \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{43.38}{z^3 - 1.838 z^2 + 0.8442 z + 0.02375}}
		\label{eq:lems}
\end{align}
\end{samepage}

\begin{samepage}
It is still necessary to address the delta inputs and outputs.
It can be seen in Figure \ref{fig:nodelta} that there is a beneficial
side effect which makes this trivial.
The transform that converts from steady state to a delta cancels
with the transform that converts from a delta to steady state
\footnote{In the more general case, when torque is non zero,
the transforms would not cancel.}.
Therefore the final transform is still Equation \ref{eq:lems}.
An example of the output response with no control is shown in
Figure \ref{fig:em_plot}.
\end{samepage}

@@

Applying this equation results in the response shown
in Figure \ref{fig:rpmtime_plot}.
It can be seen that this plot reveals that it taks over two seconds
for the system to stabilize with no control.
Clearly this is unacceptably slow and it is likely that with
the application of a controller its performance can be improved.

@@

@@

@@

>>>>

In order to build a controller the model needs to be simplified
but the presence of two inputs complicates matters.
To resolve this issue the torque can be set zero.
Then it can be simplified by recognizing that it matches
the well known form shown in Figure \ref{fig:dd} which
has the transfer function in Equation \ref{eq:dd}.

@@

\begin{samepage}
Then substituting them in to Equation \ref{eq:dd} and simplifying
results in Equation \ref{eq:lems}.
\begin{align}
	\frac{y}{u} &= \frac{(1.699) (8.5683) (2.98) (1)}
		{(z^2 - 0.9025z)(z - 0.9354)(1) + (8.5683)(2.98)(0.00093)} \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{43.38}{z^3 - 1.838 z^2 + 0.8442 z + 0.02375}}
		\label{eq:lems}
\end{align}
\end{samepage}

\begin{samepage}
It is still necessary to address the delta inputs and outputs.
This can be resolved by placing the conversion transforms,
from Figure \ref{fig:sdtf} and \ref{fig:dstf}, on either
end of the engine model transfer function as shown in Figure \ref{fig:nodelta}.
A beneficial side effect becomes apparent in this form.
The transform that converts from steady state to a delta cancels
with the transform that converts from a delta to steady state
\footnote{In the more general case, when torque is non zero,
the transforms would not cancel.}.
Therefore the final transform is still Equation \ref{eq:lems}.
An example of the output response with no control is shown in
Figure \ref{fig:em_plot}.
\end{samepage}

@@

Applying this equation results in the response shown
in Figure \ref{fig:rpmtime_plot}.
It can be seen it takes over two seconds for the system to
stabilize with no control.
Clearly this is unacceptably slow and performance could be
improved by using a controller.

@@

There is yet another complication with this model.
Typical systems are in a normalized so that it is possible to
achieve an output of one given a step input.
With this engine model an output of 1 rpm is an impossibly slow value.
And an input of zero or one rpm are similarly impossible.

To resolve this issue the values will be scaled in to a normal range.
A normal idle rpm range is from 600 to 1200 rpm.
To scale these values so that zero equals 600 and one equals 1000
Equation \ref{eq:scale} can be used.

\begin{align}
	y &= \frac{1}{400}x - \frac{3}{2} \label{eq:scale}
\end{align}

Figure \ref{fig:stable} shows the stable rpms for a given control input.
Its behavior is described by Equation \ref{eq:stable}.

\begin{align}
	y &= 1450.1 x \label{eq:stable}
\end{align}

% {{{ fig:stable
\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.6]{../octave/stable_rpm_plot}
\end{center}
\caption{Stable rpm reached for a given constant control input.}
\label{fig:stable}
\end{figure}
% }}}

With these normalization corrections applied the response can be
seen in Figure \ref{fig:normeng}.
To avoid the zero rpm case it is allowed to stabilize before the
step is applied.
The lowest plot shows the response after the step.
It takes slightly over one second to stabilize, which is better than
the inaccurate case from zero rpm, but it is still slow and can
be improved.

% {{{ fig:normeng
\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{../octave/norm_engine_model_plot}
\includegraphics[scale=0.55]{../octave/norm_engine_plot_close}
\end{center}
\caption{Normalized engine model response.  After it stabilizes
at unit step is applied to produce the normalized response.
The lowest plot is a close up of the step response in the center
plot.}
\label{fig:normeng}
\end{figure}
% }}}

@@

\clearpage
\section{Conclusion}

@@

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of the normalized engine model.}
	]{../octave/norm_engine_model_plot.m}

\clearpage
\lstinputlisting[
	caption={Matlab source to plot the response of the normalized engine model
	after step is applied.}
	]{../octave/norm_engine_plot_close.m}
